---
title: "Ethics Project"
author: "Nicholas Paisley"
date: "11/27/2020"
output: html_document
---

## This is a project based on population, police records and some basic statitiscs. 
Question: Were people of color given more citations based on Red and/or Blue states (according to the town/state population) during a time with a Blue and Red President. 

## 2012 Red States = ID UT AZ MT WY ND SD NE KS OK TX MO AR LA IN KY TN MS AL GA SC WV AK
## 2012 Blue States = WA OR CA NV CO NM MN IA WI IL MI OH FL ME VT NH MA CT RI NJ DE MD NY PA VA NC FL

## 2016 Red States = AZ ID UT MT WY ND SD NE KS OK TX IA MO AR LA WI MI IN KY TN MS AL OH GA FL WV PA NC SC AK
## 2016 Blue States = WA OR CA NV CO NM MN IL NY VA VT NH ME MA RI CT NJ DE MD HI 

##Note: Not all of these states are going to be used because some states either do not have the proper data or no data at all.

#The states that have the data needed are the following : TX, AZ, LA, FL, PA, WA, IL, NV, VT, CA
```{r}
getwd()
```

```{r, package install}
install.packages('tidyverse', dependencies = TRUE)
library(tidyverse)
library(stringr)
library(ggplot2)
library(lubridate)
```

```{r, Texas, San Antonio}
Texas_SA <- read_csv("tx_san_antonio_2020_04_01.csv")
class(Texas_SA)
```

```{r}
glimpse(Texas_SA)
```

```{r}
TXSA_2013 <- Texas_SA %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
TXSA_2017 <- Texas_SA %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(TXSA_2013)
```

```{r}
glimpse(TXSA_2017)
```

```{r}
unique(TXSA_2013$subject_race)
```

```{r}
unique(TXSA_2017$subject_race)
```

```{r}
TXSA_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
TXSA_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
TXSA_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
TXSA_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
TXSA_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
TXSA_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
TXSA_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
TXSA_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(TXSA_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(TXSA_2017$subject_race), las=2) #las changes the orientation of the tick marks
```
##Using datausa.io for population/census data
```{r}
TXSA_2013B <- (1359033 * (6.76/100)) / 18411
TXSA_2013B
TXSA_2013H <- (1359033 * (62.97705/100)) / 86693
TXSA_2013H
TXSA_2013API <- (1359033 *(2.64632/100)) / 1972
TXSA_2013API
TXSA_2013W <- (1359033 * (26.1/100)) / 74143
TXSA_2013W
TXSA_2017B <- (1461623 * (6.48/100))/ 15364
TXSA_2017B
TXSA_2017H <- (1461623 *(64.1219/100)) / 73016
TXSA_2017H
TXSA_2017API <- (1461623 * (2.52792/100)) / 1529
TXSA_2017API
TXSA_2017W <- (1461623 * (24.8/100)) / 54726
TXSA_2017W
```

```{r}
TXSA_POP_PER <- c(TXSA_2013B, TXSA_2017B, TXSA_2013API, TXSA_2017API, TXSA_2013W, TXSA_2017W, TXSA_2013H, TXSA_2017H)
TXSA_POP_PER_LABEL <- c("TXSA_2013B", "TXSA_2017B", "TXSA_2013API", "TXSA_2017API", "TXSA_2013W", "TXSA_2017W", "TXSA_2013H", "TXSA_2017H")
par(mar=c(7,3,2,3))
barplot(TXSA_POP_PER, names.arg = TXSA_POP_PER_LABEL, las=2)
```

```{r, Texas, Garland}
TEXAS_GL <- read_csv("tx_garland_2020_04_01.csv")
class(TEXAS_GL)
```

```{r}
glimpse(TEXAS_GL)
```

```{r}
TEXAS_GL2013 <- TEXAS_GL %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
TEXAS_GL2017 <- TEXAS_GL %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(TEXAS_GL2013)
```

```{r}
glimpse(TEXAS_GL2017)
```

```{r}
unique(TEXAS_GL2013$subject_race)
```

```{r}
unique(TEXAS_GL2017$subject_race)
```

```{r}
TEXAS_GL2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
TEXAS_GL2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
TEXAS_GL2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
TEXAS_GL2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
TEXAS_GL2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
TEXAS_GL2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
TEXAS_GL2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
TEXAS_GL2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```


```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(TEXAS_GL2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(TEXAS_GL2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
TXGL_2013B <- (230177 * (11.0/100)) / 6044
TXGL_2013B
TXGL_2013H <- (230177 * (35.7/100)) / 11975
TXGL_2013H
TXGL_2013API <- (230177 *(5.96/100)) / 1838
TXGL_2013API
TXGL_2013W <- (230177 * (16.3/100)) / 12178
TXGL_2013W
TXGL_2017B <- (235965 * (8.48/100))/ 5345
TXGL_2017B
TXGL_2017H <- (235965 * (36.3/100)) / 11601
TXGL_2017H
TXGL_2017API <- (235965 * (7.46/100)) / 2248
TXGL_2017API
TXGL_2017W <- (235965 * (32.5/100)) / 11515
TXGL_2017W
```

```{r, with other}
TXGL_POP_PER <- c(TXGL_2013B, TXGL_2017B, TXGL_2013API, TXGL_2017API, TXGL_2013W, TXGL_2017W, TXGL_2013H, TXGL_2017H)
TXGL_POP_PER_LABEL <- c("TXGL_2013B", "TXGL_2017B", "TXGL_2013API", "TXGL_2017API", "TXGL_2013W", "TXGL_2017W", "TXGL_2013H", "TXGL_2017H")
par(mar=c(7,3,2,3))
barplot(TXGL_POP_PER, names.arg = TXGL_POP_PER_LABEL, las=2)
```

```{r, without Other for Garland}
TXGL_POP_PER <- c(TXGL_2013B, TXGL_2017B, TXGL_2013API, TXGL_2017API, TXGL_2013W, TXGL_2017W, TXGL_2013H, TXGL_2017H)
TXGL_POP_PER_LABEL <- c("TXGL_2013B", "TXGL_2017B", "TXGL_2013API", "TXGL_2017API", "TXGL_2013W", "TXGL_2017W", "TXGL_2013H", "TXGL_2017H")
par(mar=c(7,3,2,3))
barplot(TXGL_POP_PER, names.arg = TXGL_POP_PER_LABEL, las=2)
```

```{r, Louisana, New Orleans}
LANO <- readRDS("yg821jf8611_la_new_orleans_2020_04_01.rds", refhook = NULL)
class(LANO)
```

```{r}
glimpse(LANO)
```

```{r}
LANO_2013 <- LANO %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
LANO_2017 <- LANO %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(LANO_2013)
```

```{r}
glimpse(LANO_2017)
```

```{r}
unique(LANO_2013$subject_race)
```

```{r}
unique(LANO_2017$subject_race)
```

```{r}
LANO_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
LANO_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
LANO_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
LANO_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
LANO_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
LANO_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
LANO_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
LANO_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(LANO_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(LANO_2017$subject_race), las=2) #las changes the orientation of the tick marks
```

##Using datausa.io for population/census data
```{r}
LANO_2013B <- (378623 * (58.9/100)) / 11440
LANO_2013B
LANO_2013H <- (378623 * (5.4956/100)) / 456
LANO_2013H
LANO_2013API <- (378623 *(3.0282/100)) / 167
LANO_2013API
LANO_2013W <- (378623 * (30.9/100)) / 4126
LANO_2013W
LANO_2017B <- (391538 * (58.5/100))/ 10988
LANO_2017B
LANO_2017H <- (391538 * (5.6698/100)) / 624
LANO_2017H
LANO_2017API <- (391538 * (2.91712/100)) / 122
LANO_2017API
LANO_2017W <- (391538 * (30.6/100)) / 4141
LANO_2017W
```

```{r}
LANO_POP_PER <- c(LANO_2013B, LANO_2017B, LANO_2013API, LANO_2017API, LANO_2013W, LANO_2017W, LANO_2013H, LANO_2017H)
LANO_POP_PER_LABEL <- c("LANO_2013B", "LANO_2017B", "LANO_2013API", "LANO_2017API", "LANO_2013W", "LANO_2017W", "LANO_2013H", "LANO_2017H")
par(mar=c(7,3,2,3))
barplot(LANO_POP_PER, names.arg = LANO_POP_PER_LABEL, las=2)
```

```{r, Pennsylvania, Pittsburgh}
PAPI <- readRDS("yg821jf8611_pa_pittsburgh_2020_04_01.rds", refhook = NULL)
class(PAPI)
```

```{r}
glimpse(PAPI)
```

```{r}
PAPI_2013 <- PAPI %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
PAPI_2017 <- PAPI %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(PAPI_2013)
```

```{r}
glimpse(PAPI_2017)
```

```{r}
unique(PAPI_2013$subject_race)
```

```{r}
unique(PAPI_2017$subject_race)
```

```{r}
PAPI_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
PAPI_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
PAPI_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
PAPI_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
PAPI_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
PAPI_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
PAPI_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
PAPI_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(PAPI_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(PAPI_2017$subject_race), las=2) #las changes the orientation of the tick marks
```

##Using datausa.io for population/census data
```{r}
PAPI_2013B <- (1536704 * (41.8/100)) / 4082
PAPI_2013B
PAPI_2013H <- (1536704 * (13.264/100)) / 111
PAPI_2013H
PAPI_2013API <- (1536704 *(6.73/100)) / 223
PAPI_2013API
PAPI_2013W <- (1536704 * (36.2/100)) / 7767
PAPI_2013W
PAPI_2017B <- (1569657 * (40.9/100))/ 3840
PAPI_2017B
PAPI_2017H <- (1569657 * (14.8118/100)) / 82
PAPI_2017H
PAPI_2017API <- (1569657 * (7.2757/100)) / 202
PAPI_2017API
PAPI_2017W <- (1569657 * (34.5/100)) / 4839
PAPI_2017W
```

```{r}
PAPI_POP_PER <- c(PAPI_2013B, PAPI_2017B, PAPI_2013API, PAPI_2017API, PAPI_2013W, PAPI_2017W, PAPI_2013H, PAPI_2017H)
PAPI_POP_PER_LABEL <- c("PAPI_2013B", "PAPI_2017B", "PAPI_2013API", "PAPI_2017API", "PAPI_2013W", "PAPI_2017W", "PAPI_2013H", "PAPI_2017H")
par(mar=c(7,3,2,3))
barplot(PAPI_POP_PER, names.arg = PAPI_POP_PER_LABEL, las=2)
```

```{r, California, Bakersfield}
CA_BF <- read_csv("ca_bakersfield_2020_04_01.csv")
class(CA_BF)
```

```{r}
glimpse(CA_BF)
```

```{r}
CA_BF_2013 <- CA_BF %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
CA_BF_2017 <- CA_BF %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(CA_BF_2013)
```

```{r}
glimpse(CA_BF_2017)
```

```{r}
unique(CA_BF_2013$subject_race)
```

```{r}
unique(CA_BF_2017$subject_race)
```

```{r}
CA_BF_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
CA_BF_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
CA_BF_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
CA_BF_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
CA_BF_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
CA_BF_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
CA_BF_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
CA_BF_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(CA_BF_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(CA_BF_2017$subject_race), las=2) #las changes the orientation of the tick marks
```

##Using datausa.io for population/census data
```{r}
CA_BF_2013B <- (363449 * (7.76/100)) /  1485
CA_BF_2013B
CA_BF_2013H <- (363449 * (46.561/100)) / 3353
CA_BF_2013H
CA_BF_2013API <- (363449 *(7.71/100)) / 286
CA_BF_2013API
CA_BF_2013W <- (363449 * (35.9/100)) / 6058
CA_BF_2013W
CA_BF_2017B <- (379171 * (7.35/100))/ 1903
CA_BF_2017B
CA_BF_2017H <- (379171 * (50.9765/100)) / 8126
CA_BF_2017H
CA_BF_2017API <- (379171 * (7.343/100)) / 367
CA_BF_2017API
CA_BF_2017W <- (379171 * (31.6/100)) / 6932
CA_BF_2017W
```

```{r}
CA_BF_POP_PER <- c(CA_BF_2013B, CA_BF_2017B, CA_BF_2013API, CA_BF_2017API, CA_BF_2013W, CA_BF_2017W, CA_BF_2013H, CA_BF_2017H)
CA_BF_POP_PER_LABEL <- c("CA_BF_2013B", "CA_BF_2017B", "CA_BF_2013API", "CA_BF_2017API", "CA_BF_2013W", "CA_BF_2017W", "CA_BF_2013H", "CA_BF_2017H")
par(mar=c(7,3,2,3))
barplot(CA_BF_POP_PER, names.arg = CA_BF_POP_PER_LABEL, las=2)
```

```{r, Nevada, Henderson}
NV_HD <- readRDS("yg821jf8611_nv_henderson_2020_04_01.rds", refhook = NULL)
class(NV_HD)
```

```{r}
glimpse(NV_HD)
```

```{r}
NV_HD_2013 <- NV_HD %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
NV_HD_2017 <- NV_HD %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(NV_HD_2013)
```

```{r}
glimpse(NV_HD_2017)
```

```{r}
unique(NV_HD_2013$subject_race)
```

```{r}
unique(NV_HD_2017$subject_race)
```

```{r}
NV_HD_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
NV_HD_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
NV_HD_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
NV_HD_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
NV_HD_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
NV_HD_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
NV_HD_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
NV_HD_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(NV_HD_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(NV_HD_2017$subject_race), las=2) #las changes the orientation of the tick marks
```
##Using datausa.io for population/census data
```{r}
NV_HD_2013B <- (268450 * (5.7/100)) / 1575
NV_HD_2013B
NV_HD_2013H <- (268450 * (12.0155/100)) / 2561
NV_HD_2013H
NV_HD_2013API <- (268450 *(9.183/100)) / 1024
NV_HD_2013API
NV_HD_2013W <- (268450 * (69.5/100)) / 13130
NV_HD_2013W
NV_HD_2017B <- (299631 * (6.14/100))/ 1892
NV_HD_2017B
NV_HD_2017H <- (299631 * (14.664/100)) / 1557
NV_HD_2017H
NV_HD_2017API <- (299631 * (10.732/100)) / 735
NV_HD_2017API
NV_HD_2017W <- (299631 * (64.9/100)) / 11730
NV_HD_2017W
```

```{r}
NV_HD_POP_PER <- c(NV_HD_2013B, NV_HD_2017B, NV_HD_2013API, NV_HD_2017API, NV_HD_2013W, NV_HD_2017W, NV_HD_2013H, NV_HD_2017H)
NV_HD_POP_PER_LABEL <- c("NV_HD_2013B", "NV_HD_2017B", "NV_HD_2013API", "NV_HD_2017API", "NV_HD_2013W", "NV_HD_2017W", "NV_HD_2013H", "NV_HD_2017H")
par(mar=c(7,3,2,3))
barplot(NV_HD_POP_PER, names.arg = NV_HD_POP_PER_LABEL, las=2)
```

```{r, Vermont, Burlington}
VT_BU <- readRDS("yg821jf8611_vt_burlington_2020_04_01.rds", refhook = NULL)
class(VT_BU)
```

```{r}
glimpse(VT_BU)
```

```{r}
VT_BU_2013 <- VT_BU %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
VT_BU_2017 <- VT_BU %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(VT_BU_2013)
```

```{r}
glimpse(VT_BU_2017)
```

```{r}
unique(VT_BU_2013$subject_race)
```

```{r}
unique(VT_BU_2017$subject_race)
```

```{r}
VT_BU_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
VT_BU_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
VT_BU_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
VT_BU_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
VT_BU_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
VT_BU_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
VT_BU_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
VT_BU_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(VT_BU_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(VT_BU_2017$subject_race), las=2) #las changes the orientation of the tick marks
```

##Using datausa.io for population/census data
```{r}
VT_BU_2013B <- (42671 * (3.15/100)) / 119
VT_BU_2013B
VT_BU_2013H <- (42671 * (2.4819/100)) / 15
VT_BU_2013H
VT_BU_2013API <- (42671 *(5.2089/100)) / 49
VT_BU_2013API
VT_BU_2013W <- (42671 * (86.3/100)) / 1774
VT_BU_2013W
VT_BU_2017B <- (42364 * (5.47/100))/ 93
VT_BU_2017B
VT_BU_2017H <- (42364  * (2.85907/100)) / 17
VT_BU_2017H
VT_BU_2017API <- (42364  * (6.2736/100)) / 40
VT_BU_2017API
#VT_BU_2017O <- (42364  * (.158/100)) / 0
VT_BU_2017O <- 0
VT_BU_2017O
VT_BU_2017W <- (42364 * (82.7/100)) / 581
VT_BU_2017W
```

```{r}
VT_BU_POP_PER <- c(VT_BU_2013B, VT_BU_2017B, VT_BU_2013API, VT_BU_2017API, VT_BU_2013W, VT_BU_2017W, VT_BU_2013H, VT_BU_2017H)
VT_BU_POP_PER_LABEL <- c("VT_BU_2013B", "VT_BU_2017B", "VT_BU_2013API", "VT_BU_2017API", "VT_BU_2013W", "VT_BU_2017W", "VT_BU_2013H", "VT_BU_2017H")
par(mar=c(7,3,2,3))
barplot(VT_BU_POP_PER, names.arg = VT_BU_POP_PER_LABEL, las=2)
```

```{r, California, Long Beach}
CA_LB <- read_csv("ca_long_beach_2020_04_01.csv")
class(CA_LB)
```

```{r}
glimpse(CA_LB)
```

```{r}
CA_LB_2013 <- CA_LB %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
CA_LB_2017 <- CA_LB %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(CA_LB_2013)
```

```{r}
glimpse(CA_LB_2017)
```

```{r}
unique(CA_LB_2013$subject_race)
```

```{r}
unique(CA_LB_2017$subject_race)
```

```{r}
CA_LB_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
CA_LB_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
CA_LB_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
CA_LB_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
CA_LB_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
CA_LB_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
CA_LB_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
CA_LB_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(CA_LB_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(CA_LB_2017$subject_race), las=2) #las changes the orientation of the tick marks
```

##Using datausa.io for population/census data
```{r}
CA_LB_2013B <- (468303 * (12.9/100)) / 8850
CA_LB_2013B
CA_LB_2013H <- (468303 * (42.0979/100)) / 14214
CA_LB_2013H
CA_LB_2013API <- (468303 *(13.22/100)) / 1030
CA_LB_2013API
CA_LB_2013W <- (468303 * (28.7/100)) / 8860
CA_LB_2013W
CA_LB_2017B <- (467367 * (12.3/100))/ 5142
CA_LB_2017B
CA_LB_2017H <- (467367  * (43.141/100)) / 8428
CA_LB_2017H
CA_LB_2017API <- (467367  * (13.9/100)) / 574
CA_LB_2017API
CA_LB_2017W <- (467367 * (28.1/100)) / 4185
CA_LB_2017W
```

```{r}
CA_LB_POP_PER <- c(CA_LB_2013B, CA_LB_2017B, CA_LB_2013API, CA_LB_2017API, CA_LB_2013W, CA_LB_2017W, CA_LB_2013H, CA_LB_2017H)
CA_LB_POP_PER_LABEL <- c("CA_LB_2013B", "CA_LB_2017B", "CA_LB_2013API", "CA_LB_2017API", "CA_LB_2013W", "CA_LB_2017W", "CA_LB_2013H", "CA_LB_2017H")
par(mar=c(7,3,2,3))
barplot(CA_LB_POP_PER, names.arg = CA_LB_POP_PER_LABEL, las=2)
```

```{r, Florida, Tampa}
FL_TP <- read_csv("fl_tampa_2020_04_01.csv")
class(FL_TP)
```

```{r}
glimpse(FL_TP)
```

```{r}
FL_TP_2013 <- FL_TP %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2013-01-01"), as.Date("2013-12-31")))
FL_TP_2017 <- FL_TP %>% select(date, subject_race, citation_issued) %>% filter(between(date, as.Date("2017-01-01"), as.Date("2017-12-31")))
```

```{r}
glimpse(FL_TP_2013)
```

```{r}
glimpse(FL_TP_2017)
```

```{r}
unique(FL_TP_2013$subject_race)
```

```{r}
unique(FL_TP_2017$subject_race)
```

```{r}
FL_TP_2013 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
FL_TP_2013 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
FL_TP_2013 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
FL_TP_2013 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
FL_TP_2017 %>% filter(subject_race == "black" , citation_issued == TRUE) 
```

```{r}
FL_TP_2017 %>% filter(subject_race == "hispanic" , citation_issued == TRUE) 
```

```{r}
FL_TP_2017 %>% filter(subject_race == "asian/pacific islander" , citation_issued == TRUE) 
```

```{r}
FL_TP_2017 %>% filter(subject_race == "white" , citation_issued == TRUE) 
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(FL_TP_2013$subject_race), las=2) #las changes the orientation of the tick marks
```

```{r}
par(mar=c(8.75,4,2,1)) #changes the margins of the graph (bottom, left, top, right)
barplot(table(FL_TP_2017$subject_race), las=2) #las changes the orientation of the tick marks
```
##Using datausa.io for population/census data
```{r}
FL_TP_2013B <- (352426 * (23.5/100)) / 55318
FL_TP_2013B
FL_TP_2013H <- (352426 * (23.782/100)) / 14147
FL_TP_2013H
FL_TP_2013API <- (352426 *(4.1442/100)) / 1765
FL_TP_2013API
FL_TP_2013W <- (352426 * (45.9/100)) / 134912
FL_TP_2013W
FL_TP_2017B <- (386577 * (22.1/100))/ 29984
FL_TP_2017B
FL_TP_2017H <- (386577 *(26.44/100)) / 16485
FL_TP_2017H
FL_TP_2017API <- (386577 * (4.2818/100)) / 2654
FL_TP_2017API
FL_TP_2017W <- (386577 * (44.7/100)) / 62745
FL_TP_2017W
```

```{r}
FL_TP_POP_PER <- c(FL_TP_2013B, FL_TP_2017B, FL_TP_2013API, FL_TP_2017API, FL_TP_2013W, FL_TP_2017W, FL_TP_2013H, FL_TP_2017H)
FL_TP_POP_PER_LABEL <- c("FL_TP_2013B", "FL_TP_2017B", "FL_TP_2013API", "FL_TP_2017API", "FL_TP_2013W", "FL_TP_2017W", "FL_TP_2013H", "FL_TP_2017H")
par(mar=c(8,3,2,4))
barplot(FL_TP_POP_PER, names.arg = FL_TP_POP_PER_LABEL, las=2)
```

```{r}
par_col <- c("black" , "goldenrod", "green", "blue")
```

 
```{r, Creates graph with variables }
Y2013 <- c(TXSA_2013B, TXSA_2013H, TXSA_2013API, TXSA_2013W) #numerical variables
Y2017 <- c(TXSA_2017B, TXSA_2017H, TXSA_2017API, TXSA_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Red to Red: San Antonia, Texas", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```

```{r}
Y2013 <- c(TXGL_2013B, TXGL_2013H, TXGL_2013API, TXGL_2013W) #numerical variables
Y2017 <- c(TXGL_2017B, TXGL_2017H, TXGL_2017API, TXGL_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Red to Red: Garland, Texas", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```

```{r}
Y2013 <- c(LANO_2013B, LANO_2013H, LANO_2013API, LANO_2013W) #numerical variables
Y2017 <- c(LANO_2017B, LANO_2017H, LANO_2017API, LANO_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Red to Red: New Orleans, Louisiana", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```

```{r}
Y2013 <- c(PAPI_2013B, PAPI_2013H, PAPI_2013API, PAPI_2013W) #numerical variables
Y2017 <- c(PAPI_2017B, PAPI_2017H, PAPI_2017API, PAPI_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Blue to Red: Pittsburgh, Pennsylvania", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```

```{r}
Y2013 <- c(NV_HD_2013B, NV_HD_2013H, NV_HD_2013API, NV_HD_2013W) #numerical variables
Y2017 <- c(NV_HD_2017B, NV_HD_2017H, NV_HD_2017API, NV_HD_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Blue to Blue: Henderson, Nevada", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```

```{r}
Y2013 <- c(VT_BU_2013B, VT_BU_2013H, VT_BU_2013API, VT_BU_2013W) #numerical variables
Y2017 <- c(VT_BU_2017B, VT_BU_2017H, VT_BU_2017API, VT_BU_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Blue to Blue: Burlington, Vermont", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```


```{r}
Y2013 <- c(CA_LB_2013B, CA_LB_2013H, CA_LB_2013API, CA_LB_2013W) #numerical variables
Y2017 <- c(CA_LB_2017B, CA_LB_2017H, CA_LB_2017API, CA_LB_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Blue to Blue: Long Beach, California", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```

```{r}
Y2013 <- c(CA_BF_2013B, CA_BF_2013H, CA_BF_2013API, CA_BF_2013W) #numerical variables
Y2017 <- c(CA_BF_2017B, CA_BF_2017H, CA_BF_2017API, CA_BF_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Blue to Blue: Bakersfield, California", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```

```{r}
Y2013 <- c(FL_TP_2013B, FL_TP_2013H, FL_TP_2013API, FL_TP_2013W) #numerical variables
Y2017 <- c(FL_TP_2017B, FL_TP_2017H, FL_TP_2017API, FL_TP_2017W) #numerical variables
z <- c(0,0,0,0,1,1,1,1) #left=0, right=1 (x variables)
par(mar=c(3,8,1,8)) #changes the margins of the graph (bottom, left, top, right)
plot(z, c(Y2013, Y2017), xaxt="n", ylab="Average", xlab="", xlim=c(-0.5,1.5), main = "Blue to Red: Tampa, Florida", col=1, pch = c("B", "H", "API", "W"))
axis(side=1, at=c(0,1), c("2013", "2017"))
segments(rep(0,4), Y2013, rep(1,4), Y2017, col=par_col, lwd =2)
```