---
title: "R Notebook"
output: html_notebook
---


```{r}
data1 <- read.csv("TrainData_Group1.csv", header = TRUE)
```

```{r}
str(data1)
```

```{r}
summary(data1)
```


```{r}
plot(data1$X1,data1$Y)
```

```{r}
df = subset(data1, select = c("X1","X2", "X3","X4","X5"))
```

y=βX+ϵ

where ‘y’ is a vector of the response variable, ‘X’ is the matrix of our feature variables (sometimes called the ‘design’ matrix), and β is a vector of parameters that we want to estimate. ϵ is the error term; it represents features that affect the response, but are not explicitly included in our model.

```{r}
x <- as.matrix(data1[-ncol(data1)])
y <- data1$Y
summary(x)
summary(y)
```

```{r}
int <- rep(1, length(y))
summary(int)
```

```{r}
x <- cbind(int,x)
```

```{r}
B <- solve(t(x) %*% x) %*% t(x) %*% y
```


```{r}
B <- round(B,2)
print(B)
```
```{r}
# Linear regression model
lm.mod <- lm(y ~ ., data=df)

# Round for easier viewing
lm.B <- round(lm.mod$coefficients, 2)

# Create data.frame of results
results <- data.frame(our.results=B, lm.results=lm.B)

print(results)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
